Emulsions Stabilized by Magnetic Ellipsoidal Particles: A Lattice Boltzmann Study

### 2025-04-05 chapters/paper1_new.tex
Changes 1. Rewritten literature review and introduction to combine both 2. Beginning to rewrite methods section 3. Finished rewrite of aim 1 - magnetic fields 4. Finished adding structure characterization

This section is based on work published in the Royal Society of Chemistry's Soft Matter journal and
work submited to the American Institude of Physics, Physics of Fluids journal. \cite{karthikeyan_formation_2024} 
It is reproduced from Ref \cite{karthikeyan_formation_2024} with permission from the Royal Society of Chemistry and 
reproduced from submission number \#POF25-AR-DSFD2024-03180, with the permission of AIP Publishing.

The addition of microstructure control in bijels can be utilized to improve the performance of porous materials
or soft matter derived from their co-continuous, tortuous microstructure. The importance of microstructure control
allows an increase in the charge/discharge rate of batteries through reducing the material tortuosity, or improving the 
adhesion and growth rate of bone grafts. These properties are imparted during the fabrication step of the bijel. Anisotropic
particles adsorbed at interfaces induce shape-dependent capillary interactions that have distinct effects on the
microstructure of emulsions, resulting in migration or rotation of particles at interfaces, in addition to the addition of
distinct timescales of domain coarsening. \cite{loudet_capillary_2005, cavallaro_curvature-driven_2011, gunther_lattice_2013}

In bijels, ellipsoidal stabilizers have been demonstrates to result in additional timescales of domain coarsening 
at long timescales owing to particle re-orientation at the interface due to interparticle capillary interactions 
at long timescales. \cite{gunther_timescales_2014} In addition to possessing more timescales, property, bijels stabilized 
with rod-like particles have been shown to result in smaller characteristic length scales, owing to the particles having a
larger cross sectional area resulting in jamming occurring at larger interfacial areas. \cite{hijnen_bijels_2015}

Magnetic response has been used in particle stabilized emulsions in the past, causing controlled emulsion failure and
on-demand droplet coalescence for encapsulation and cargo release applications. \cite{melle_pickering_2005, zhou_magnetic_2011}
Fields also enable self assembly through the control of interface deformation derived capillary forces.
\cite{morgan_understanding_2013,davies_interface_2014,davies_dipolar_2015} Magnetically responsive spherical particle 
stabilized bijels have thus far not yielded significant microstructural modifications. \cite{kim_bijels_2010} However, 
electric fields have been shown to cause percolating cylindrical domains in electrically responsive bijels. \cite{carmack_tuning_2018}
The electric field causes polar interactions between particles, which causes the formation of chains of particles that
also function as nucleation sites for phase separation. \cite{carmack_tuning_2018}

In this chapter, We investigate the effect of external magnetic fields on the formation of bijels stabilized by 
anisotropic magnetic particles. The microstructure of bijels stabilized with oblate, spherical and prolate particles, 
namely the characteristic length scale and tortuosity, will be assessed as a function of the applied magnetic field
strength. We also characterize the underlying mechanisms governing the observed behavior.

\section{Results}\label{sec:results_p1}

Bijels serve as emulsion templates for fabricating porous materials with applications in areas such as drug delivery, water purification, and energy 
storage devices like battery electrodes \cite{vanoli_bijels_2022, chen_pore-scale_2022, lu_controllable_2020, garcia_scalable_2019}. The performance of
materials in these applications are influenced by their microstructure. In the following section, the microstructure of magnetically responsive bijels
are assessed. The bijels in our simulations have a particle volume fraction $\phi_p = 0.15$ with magnetic Bond numbers ranging from 
$\bar{Bo} = 0, 0.2, 0.5, 1$. Three simulations per particle and per magnetic field strength are performed, allowing for investigations into the effect of
the initial particle configuration on the obtained microstructure. While properties such as the particle contact angle, fluid ratio, surface tension and
particle volume fraction are important, this study will focus exclusively on the effect of magnetic fields on bijels. 
\cite{jansen_bijels_2011, hijnen_bijels_2015}

\subsection{Microstructure evolution of magnetically responsive bijels}

\begin{figure}
    \centering
    \includegraphics[width=0.5\columnwidth]{figures/results/paper1/microstructure_viz.png}
    \caption{Snapshots of emulsion gels stabilized by prolate ellipsoids after $t=5\cdot10^3$ timesteps (left) and after $10^5$ timesteps (right). The top 
             row shows the structure forming without a magnetic field. The bottom row shows the structure forming in an applied magnetic field of reduced 
             strength $\bar{B}=1.0$. The snapshots are colored according to the order parameter $\phi$.}
    \label{fig:microstructure_viz}
\end{figure}

Figure \ref{fig:microstructure_viz} presents snapshots of the bijel microstructure during formation and at the final timestep with and without magnetic fields.
The figure highlights the alignment of anisotropic particles along the field direction following its application. Particle reorientation leads to apparently larger
domain sizes in the bijels, characterized through magnetic field driven rearrangement of particles at the interface. Past work has identified how droplet coalescence
can be induced through application of magnetic fields. \cite{melle_pickering_2005} This behavior can be quantified using the domain size, done in the following section.

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/results/paper1/domain_size.png}
    \caption{Time dependence of the average domain size of the bijel for different particles shapes $\alpha$ and at different magnetic 
             field strength $\bar{B}$. $L_1$ denotes the domain size obtained from the first moment of the spherical average structure 
             factor, and $L_d$ denotes the domain size obtained from the second moment of the 3D structure factor. The increase roughly 
             follows a $\sim t^{2/3}$ scaling law indicative of the inertial regime of spinodal decomposition. Image courtesy of Ulf D. Schiller}
    \label{fig:domain_size}
\end{figure}

Two methods have been identified to characterize the domain size of bijels from the structure factor. They are the first moment of the spherically averaged structure factor
named $L_1$ and the second moment of the directional structure factor. The numerical average of the second moment derived structure factor is named $L_d$. Comparing the time
evolution of both in Figure \ref{fig:domain_size}, the domain size increases before slowing down and finally plateauing. This indicates unhindered spinodal decomposition at
early times, followed by a slowdown of domain coarsening caused by particle interactions before the bijel microstructure jams. Comparing the characterized domain sizes, 
$L_d \~ 80$ compared to $L_1 \~ 150$. $L_d$ and $L_1$ also have differing time evolution, with $L_d$ beginning to plateau earlier than $L_1$. To identify which lengthscale 
should be used to characterize the time evolution of the structure, we invoke dynamical scaling.

From dynamical scaling, spinodal decomposition of binary fluids can be characterized through the inertial regime $L \~ t^{2/3}$, crossover regime $L \~ t^{n}, 2/3 < n < 1$ and 
the inertial regime $L \~ t$. \cite{kendon_3d_1999,kendon_inertial_2001} These scaling regimes can be obtained from the Navier Stokes equation with additional surface tension 
forces. The characterized lengthscales should conform to dynamical scaling as it is independent of the method used to characterize the lengthscale. By fitting the first phase
of the domain size evolution for all particle shapes, the scaling exponent of domain size change characterized with $L_1$ is found to be approximately $L \~ t^{2/3}$. However
the scaling exponent for $L_d$ is found to be $L \~ t^{0.2}$. Therefore, $L_1$ is selected to characterize the time evolution of the domain size. However the second moment of
the domain size allows characterization of the domain size in each cartesian direction. The second moment derived domain size will be used to characterize the microstructure
of the bijel.

\subsection{Effect of magnetic field on bijel microstructure}

% In LB simulations of spinodal decomposition, a characteristic length
% scale can be obtained from the structure factor, defined in Section \ref{section:domain_size}. \cite{kendon_3d_1999,kendon_inertial_2001} 
% It is worth noting that this definition of average domain size does not necessarily yield values that coincide with the typical
% domain size observed visually in snapshots such as those shown in Figure~\ref{fig:microstructure_viz}. However, in a dynamic
% scaling regime, different measures of domain size are expected to differ only by a constant ratio. We found that the commonly 
% used measure \(L_1(t)\) tends to overestimate the observed domain size. Numerically, the calculation of the spherically 
% averaged structure factor is subject to poor statistics in the low \(k\)-shells, where the average \(|\vec{k}|\) of the 
% lattice sites differs from the nominal shell radius. Additionally, the measurement of \(L(t)\) in simulations with
% periodic boundary conditions is subject to finite size effects. More importantly, in suspensions of anisotropic particles, 
% We cannot expect that the structure factor is isotropic. Therefore, we also used the 3D structure factor to calculate a lateral domain 
% size in each Cartesian direction from the second moments also defined in Section \ref{section:domain_size}. 
% \cite{jansen_bijels_2011,gunther_timescales_2014} We add the additional processing calculating a parallel and perpendicular domain size
% from the results in each cartesian direction, defining $L_{\parallel}=L_z$ and $L_{\perp} = \frac{L_x+L_y}{2}$ we plot the structure factor of the structures in
% Figure \ref{fig:structure_factor}.

\begin{figure}
    \centering
    \includegraphics[scale = 0.4]{figures/results/paper1/D2a-vs-B_ss.png}
    \caption{Dependence of the average domain size on the magnetic field strength $\bar{B}$. The left plot shows the average domain size $L_1$ obtained from the spherically averaged structure factor. 
    The right plot shows the parallel and perpendicular domain size obtained from the respective second moments of the 3D structure factor. Errorbars indicate the standard deviation taken over 
    three independent simulation runs. For anisotropic particles, the directional domain size becomes more anisotropic with increasing field strength.}
    \label{fig:D2a_B}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[scale = 0.4]{figures/results/paper1/tortuosity_compare.png}
    \caption{Dependence of the tortuosity on the magnetic field strength $\bar{B}$ for different particle shapes $\alpha$. 
             Different components of the tortuosity tensor show the anisotropy for oblate and prolate particles.}
    \label{fig:tau_B}
\end{figure}

\begin{figure} 
    \centering 
    \includegraphics[scale = 0.4]{figures/results/paper1/curvature-vs-B_ss.png} 
    \caption{Plot of the area averaged mean $\langle H \rangle \Sigma^{-1}$ and Gaussian 
            $\langle K \rangle \Sigma^{-2}$ curvature in the left and right respectively. Each 
            plot contains the $\langle H \rangle \Sigma^{-1}$ and $\langle K \rangle \Sigma^{-1}$ 
            at the final timestep averaged across 3 runs for particles with $\alpha = 0.5, 1, 2$. 
            The mean curvature remains 0 on average, while the Gaussian 
            curvature shows a reduction as the magnetic field strength is increased for 
            ellipsoidal particles.} 
    \label{fig:curvature-vs-B_ss}
\end{figure}

\subsection{Curvature}

Beyond the length scale, curvature analysis offers insights into the
topological landscape of the bijel, facilitating further
microstructure optimization for various applications.
\cite{reeves_quantitative_2016} We plot the area averaged mean and
gaussian curvature \(H\Sigma^{-1}\) and \(K\Sigma^{-2}\) for bijels
stabilized with oblate, spherical and prolate ellipsoids at the final timestep in Figure
\ref{fig:curvature-vs-B_ss}.

\begin{figure} 
    \centering 
    \includegraphics[scale = 0.4]{figures/results/paper1/curvature-vs-B_ss.png} 
    \caption{Plot of the area averaged mean $\langle H \rangle \Sigma^{-1}$ and Gaussian 
            $\langle K \rangle \Sigma^{-2}$ curvature in the left and right respectively. Each 
            plot contains the $\langle H \rangle \Sigma^{-1}$ and $\langle K \rangle \Sigma^{-1}$ 
            at the final timestep averaged across 3 runs for particles with $\alpha = 0.5, 1, 2$. 
            The mean curvature remains 0 on average, while the Gaussian 
            curvature shows a reduction as the magnetic field strength is increased for 
            ellipsoidal particles.} 
    \label{fig:curvature-vs-B_ss}
\end{figure}

Error bars are represented by the standard
deviation of the three simulations per parameter set performed. The area
averaged mean curvature shows little dependence on the applied field
strength with values around 0 and within error of one another. A value
near or at 0 is expected as the fluid volume fraction is equal and the
particles are neutrally wetting, implying the interface should not have
a preferred direction of curvature. \cite{jinnai_interfacial_2001} For
ellipsoidal particles, the gaussian curvature becomes more negative as
larger field strengths are applied. We also see that the gaussian
curvature becomes more positive as the interfacial area of the particle
increases. The trend of these results are inverse to the interfacial
area of each particle, \(A_{rm}\). This is expected behavior and is in
agreement with experiments that have investigated the effect of particle
size on the curvature of bijels. \cite{reeves_quantitative_2016}

To explain the reduction in $\langle K\Sigma^{-2} \rangle$ as a function of the applied field strength, two mechanisms are proposed. The first is that 
the particle alignment to the magnetic field during coarsening changes the topology of the surface, which upon jamming of ellipsoidal particles is locked in. 
The other proposed mechanism is that the time evolution of spinodal decomposition remains unaffected, but that the interface conforms to the magnetic field 
driven arrangement of particles upon jamming. While we cannot apply the Gauss Bonnet theorem to characterize the results, we can still relate the coarsening 
behavior of the bijel through its change in genus to the curvature. The time evolution of the genus serves as a topological technique to characterize the 
coarsening of the bijel, which alongside the $\langle K\Sigma^{-2} \rangle$, allows us to identify deviations from the coarsening behavior of bijels 
stabilized with spherical particles, and with magnetic fields. We calculate the genus by extracting the isosurface of the filled order parameter, where 
$\phi = 0$. It is then meshed using the marching cubes algorithm implemented in scikit-image with a cube size of 1. \cite{van2014scikit} From the vertices(v) 
and faces(f) generated from the marching cubes algorithm, we calculate the number of edges(e). The Euler Poincare characteristic, $\chi = n + f - e$. The genus is 
then obtained from the Euler Poincare parameter as $g = 1 - \chi/2$.

\begin{figure} 
    \centering 
    \includegraphics[scale = 0.4]{figures/results/paper1/genus_curvature_vs_coverage.png} 
    \caption{Plots of Gaussian $\langle K \rangle \Sigma^{-2}$ curvature over time averaged across 3 runs. 
    From left to right, We have particles with $\alpha = 0.5, 1, 2$. We observe that the initial stages of 
    evolution are unaffected by the applied magnetic field. However, after about 10000 timsteps, We see a 
    magnetic field dependence on the evolution of the Gaussian curvature. The preservation of this divergence 
    occurs when ellipsoidal particles are used but not spherical particles.} 
    \label{fig:curvature_vs_coverage}
\end{figure}

The time evolution of $g \Sigma^{-1}$ of bijels stabilized with spherical particles is unaffected by the application of a magnetic field. However, a 
slight decrease in the genus as a function of the applied field is observed for bijels stabilized with ellipsoidal particles. This suggests that the 
evolution of the surface corresponding to the interface is changed by the adsorption and orientation of magnetically responsive ellipsoidal particle. 
We see effects of this change in $\langle G \rangle \Sigma^{-2}$ where we see the evolution of the curvature being delayed, indicated as a rightward 
shift of the plots as a function of the applied field.

>>>>

Emulsions formed from mixtures of immiscible liquids are widely used as structural 
templates for fabricating porous materials. These include a broad array of applications 
such as porous polymers, ceramics, synthetic nanoparticles, and engineered tissue scaffolds 
\cite{zhang_emulsion_2019, stubenrauch_emulsion_2018, jones_high-temperature_2009, binks_macroporous_2002, 
abbadian_chaleshtari_nanoemulsion_2022, aldemir_dikici_basic_2020}. The technique of emulsion templating 
provides a flexible and efficient approach to controlling key structural features of the final
material—namely, the domain size, volume fraction, and phase composition—by manipulating the 
formulation and processing conditions of the emulsion \cite{mudassir_fundamentals_2021}. To ensure 
thermodynamic and mechanical stability during processing, emulsions are typically stabilized by additives. 
Among these, particle-stabilized systems—commonly referred to as Pickering emulsions—have emerged as attractive 
alternatives to traditional surfactant-based emulsions due to their ability to form stable interfaces without 
relying on potentially harmful chemicals \cite{ramsden_separation_1903, zheng_pickering_2022, menner_particle-stabilized_2007}.

A particularly intriguing subclass of Pickering emulsions is the bicontinuous interfacially jammed emulsion gel, or bijel. 
Bijels are characterized by a tortuous co-continuous morphology where two fluid phases are separated by a particle-laden 
interface \cite{cates_bijels_2008}. This structure forms when a binary liquid mixture containing colloidal particles is 
rapidly quenched into a spinodal decomposition regime. As phase separation proceeds, the particles are swept up by the 
growing interface. Eventually, particle crowding leads to jamming at the interface, effectively halting further coarsening 
and locking the structure into a bicontinuous state. Bijels were first observed in simulation studies \cite{stratford_colloidal_2005}, 
and later successfully fabricated in experimental systems involving water-lutidine mixtures stabilized with silica particles 
\cite{clegg_emulsification_2007, herzig_bicontinuous_2007}. Since these foundational discoveries, a wide range of bijel 
formulations and processing techniques have been developed to enable greater scalability and control. Notably, 
Haase et al.~\cite{haase_continuous_2015} introduced solvent transfer-induced phase separation (STRIPS), a scalable 
and continuous fabrication method that leverages imposed flow fields to direct droplet and jet formation. As a result 
of these innovations, bijels have gained widespread interest as structural templates for functional materials, including 
filtration membranes, electrochemical electrodes, and biocompatible scaffolds \cite{yabuno_preparation_2020, pang_highly_2020, 
witt_microstructural_2016, thorson_composite_2018}.

Beyond chemical composition, the geometry of the stabilizing particles plays a critical role in determining 
the microstructure and stability of emulsions. When anisotropic particles—such as ellipsoids or rods—adsorb at 
fluid interfaces, they give rise to shape-dependent capillary interactions that are markedly different from those 
induced by spherical particles \cite{loudet_capillary_2005,madivala_exploiting_2009}. These interactions can 
significantly enhance emulsion stability, as demonstrated by Madivala et al., and often result in distinct particle 
arrangements at the interface \cite{madivala_self-assembly_2009}. Moreover, anisotropic particles are capable of 
self-assembling in response to interface curvature gradients, where the curvature acts as an external field influencing 
particle orientation and spatial distribution \cite{cavallaro_curvature-driven_2011, furst_directing_2011}. Computational 
studies, including Monte Carlo and Lattice Boltzmann simulations, have provided insight into how such particles dynamically 
interact with fluid interfaces \cite{bresme_orientational_2007, bresme_computer_2008, gunther_lattice_2013,gunther_timescales_2014}. 
These simulations show that anisotropic particles introduce additional timescales during 
emulsion formation due to their rotational behavior. Initially, they rotate to align their major axes with the interface, 
and later reorient under capillary forces, contributing to prolonged coarsening and finer domain features. Experimental studies 
by Hijnen et al.~\cite{hijnen_bijels_2015} confirmed that rod-shaped particles can jam more quickly than spheres at smaller 
interfacial separations and can even tilt out of the plane due to compression during domain evolution.

Incorporating magnetic functionality into particle stabilizers introduces the possibility of externally controlling interfacial 
behavior using applied magnetic fields \cite{melle_pickering_2005, zhou_magnetic_2011}. Magnetic particles can align or assemble 
in response to external fields, enabling modulation of both emulsion stability and particle organization at the interface 
\cite{dassanayake_structure_2000, leunissen_directing_2009}. This is especially pronounced in systems containing anisotropic magnetic 
particles, where external fields can trigger orientation transitions and alter capillary interactions \cite{morgan_understanding_2012, 
davies_interface_2014, davies_dipolar_2015}. However, not all particle geometries respond equally. Simulations by Kim et al. revealed 
that magnetic fields exert minimal influence on the microstructure of bijels stabilized by spherical particles \cite{kim_bijels_2010}. 
In contrast, Carmack and Millett demonstrated that applying electric fields to bijels stabilized with dielectric particles can induce 
significant morphological changes, including domain alignment and cylindrical channel formation via field-induced dipolar interactions 
\cite{carmack_tuning_2018}. These results suggest that the interplay between field direction, particle geometry, and interface dynamics 
can be exploited to engineer emulsion microstructures in situ.

This chapter investigates how external magnetic fields influence the formation and microstructure of bijels stabilized by anisotropic 
magnetic particles. Using a hybrid simulation framework combining Lattice Boltzmann and Molecular Dynamics, we model the phase separation 
of binary fluids containing anisotropic particles with permanent dipole moments aligned along their symmetry axes. By systematically varying 
magnetic flux density and particle aspect ratio, we assess how these parameters affect domain structure, spacing, and tortuosity. Our 
findings reveal that, while magnetic fields do not inhibit bijel formation, they significantly impact the resulting morphology. 
Specifically, anisotropic particles subjected to magnetic torque align during formation, leading to anisotropic domain features and 
enhanced interfacial order. This reorientation improves packing efficiency and delays the onset of interfacial jamming. Compared to 
field-free cases, the resulting bijels display a higher degree of structural organization. These results underscore the potential of 
magnetic anisotropic particles as tools for designing responsive, field-tunable porous materials.

\section{Results}

The transport characteristics of Bijels are closely linked to the morphology of the underlying 
phase-separated bijel template. In this study, we investigate how domain size and tortuosity evolve in bijels stabilized by anisotropic particles, 
with a particular focus on the influence of magnetic fields during formation.
To this end, we simulate bijels composed of ellipsoidal magnetic particles at a fixed particle volume fraction of \(\phi_p = 0.15\), systematically 
varying the applied magnetic flux density such that the Bond number spans the range \(0 \le \bar{B} \le 1\). Our approach isolates the effects of 
the magnetic field on microstructure while holding other parameters constant. Prior studies have explored how changes in particle concentration and 
interfacial tension affect bijel morphology—for example, work by Hijnen et al.~\cite{hijnen_bijels_2015} and Jansen and Harting~\cite{jansen_bijels_2011}.

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../figures/results/paper1/microstructure_viz.png}
    \caption{Snapshots of emulsion gels stabilized by prolate ellipsoids after $t=5\cdot10^3$ timesteps (left) and after $10^5$ timesteps (right). 
             The top row shows the structure forming without a magnetic field. The bottom row shows the structure forming in an applied magnetic 
             field of reduced strength $\bar{B}=1.0$. The snapshots are colored according to the order parameter $\phi$.}
    \label{fig:microstructure_viz}
\end{figure}

Figure~\ref{fig:microstructure_viz} presents snapshots of bijel structures at various time points, comparing systems formed with and without an applied 
magnetic field. These images demonstrate how anisotropic particles align along the field direction, causing interfacial deformation and rearrangements 
such as domain reorientation, coalescence, or breakup. This alignment appears to result in larger characteristic domain sizes when magnetic fields are 
applied. A detailed, quantitative analysis of the impact of magnetic fields on bijel morphology—particularly domain size—is presented in the following sections.

\subsection{Structure factor and domain size of magnetically responsive bijels}

    In LB simulations of spinodal decomposition
    \cite{kendon_3d_1999,kendon_inertial_2001}, a characteristic length
    scale can be obtained from the structure factor
    %
    \begin{equation}
    S(\vec{k},t) = \tilde{\phi}(\vec{k},t)\tilde{\phi}(-\vec{k},t) .
    \end{equation}
    %
    where \(\tilde{\phi}(\vec{k},t)\) is the Fourier
    transform of the order parameter fluctuations
    \(\phi(\vec{x},t)-\left\langle\phi\right\rangle\). A common measure of
    domain size uses the spherically averaged structure factor
    %
    \begin{equation}
    S(k,t) = \frac{1}{n_k} \sum_{k-\frac{\Delta}{2}<k<k+\frac{\Delta}{2}} \tilde{\phi}(\vec{k},t)\tilde{\phi}(-\vec{k},t) ,
    \end{equation}
    %
    where \(k=|\vec{k}|\) denotes the modulus of \(\vec{k}\)
    and \(n_k\) is the number of lattice sites in a shell of radius \(k\)
    and thickness \(\Delta=2\pi/L_V\). The average domain size can then be
    defined in terms of the \(n\)-th moment of the structure factor
    \cite{laradji_molecular_1996}
    %
    \begin{equation}
        L_n(t) = 2\pi \left( \frac{\sum_k S(k,t)}{\sum_k k^n S(k,t)} \right)^{\frac{1}{n}}
    \end{equation}

The average domain size defined here may not directly match the visually estimated size seen in snapshots such as Fig.~\ref{fig:microstructure_viz}. 
However, in the dynamic scaling regime, different domain size metrics typically differ only by a constant factor. We observed that the commonly used 
measure \(L_1(t)\) tends to overestimate the apparent domain size. This is partly due to poor statistics in the low-\(k\) shells of the spherically 
averaged structure factor, where the average \(|\vec{k}|\) deviates from the nominal shell radius. Additionally, simulations with periodic boundaries 
are subject to finite-size effects. In systems with anisotropic particles, the structure factor is generally not isotropic. To account for this, we 
also computed directional domain sizes along each Cartesian axis from the 3D structure factor using second-moment analysis \cite{jansen_bijels_2011, 
gunther_timescales_2014}.

    %
    \begin{equation}
    L_\beta(t)=2\pi\sqrt{\frac{\sum_{\vec{k}}S(\vec{k},t)}{\sum_{\vec{k}}k_\beta^2 S(\vec{k},t)}} .
    \end{equation}
    %
    This allows us to compute a domain size parallel
    (\(L_\parallel=L_z\)) and perpendicular (\(L_\perp=(L_x+L_y)/2\)) to the
    direction of \(\vec{B}\) separately. The average domain size \(L_d(t)\)
    is computed as \(L_d(t)=\sum_\beta L_\beta(t) / 3\).
    %
    \begin{equation}
    L_d(t)=2\pi\sqrt{\frac{\sum_{\vec{k}}S(\vec{k},t)}{\sum_{\vec{k}}\sum_\beta k_\beta^2 S(\vec{k},t)}} .
    \end{equation}
    %

Before computing the structure factor \(S(\vec{k}, t)\), we filled lattice sites occupied by particles with the average density of 
their neighboring fluid nodes. This was done using an iterative method that applies Eq.~\eqref{eq:fill_particles}, starting from the 
particle surface and proceeding layer by layer until all solid nodes were filled. The order parameter \(\phi(\vec{x}, t)\) was then 
computed from the resulting filled density fields \(\rho_k(\vec{x}, t)\). This filling procedure is essential to eliminate spurious 
oscillations in the structure factor at length scales comparable to the particle size.

    \begin{figure*}
        \centering
        \includegraphics[width=\textwidth]{../figures/results/paper1/structure_factor.png}
        \caption{Scaling plot of the dynamic structure factor $S(k,t)$ for different particle shapes $\alpha$ and at different magnetic field 
                 strength $\bar{B}$. The structure factor is plotted at four different timesteps for each case. The good data collapse shows 
                 that the bijel formation is driven by spinodal decomposition. The decay to the right of the peak is reasonably well described 
                 by Porod's law $S(k)\sim k^{-4}$.}
    \label{fig:structure_factor}
    \end{figure*}

Figure~\ref{fig:structure_factor} shows the scaling behavior of the dynamic structure factor \(S(k, t)\). While the data exhibit good overall 
collapse, deviations from dynamic scaling appear at smaller length scales. The post-peak decay follows Porod's law, \(S(k) \sim k^{-4}\), indicating 
scattering from relatively flat interfaces. These findings are consistent with results by Kendon et al.~\cite{kendon_3d_1999, kendon_inertial_2001} 
for spinodal decomposition in binary fluids, supporting the view that bijel formation is initially governed by fluid demixing dynamics. Particle effects 
become significant only at later stages, when interface coarsening reduces available surface area and triggers particle interactions. This leads to jamming, 
which is reflected in the abrupt slowing of domain growth, as shown in Figure~\ref{fig:domain_size}.

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{../figures/results/paper1/domain_size.png}
    \caption{Time dependence of the average domain size of the bijel for different particles shapes $\alpha$ and at different magnetic field strength $\bar{B}$. $L_1$ denotes the domain size obtained from the first moment of the spherical average structure factor, and $L_d$ denotes the domain size obtained from the second moment of the 3D structure factor. The increase roughly follows a $\sim t^{2/3}$ scaling law indicative of the inertial regime of spinodal decomposition.}
    \label{fig:domain_size}
\end{figure*}

A nonlinear fit of the domain size with form \(b(t - t_0)^a\) over the time interval \(t \in [0,\ 3 \cdot 10^5 \Delta x]\) yields scaling exponents of \(a = 0.6448\) 
for spherical particles, \(a = 0.6083\) for oblate particles with aspect ratio \(\alpha = 0.5\), and \(a = 0.652\) for prolate ellipsoids with \(\alpha = 2\). 
These values are consistent with the theoretical exponent \(a = 2/3\), characteristic of the inertial scaling regime, where the characteristic velocity 
\(U = dL/dt\) scales as \(U \sim \sqrt{\sigma / (\rho L)}\), resulting in \(L \sim t^{2/3}\).
Beyond approximately 30,000–40,000 time steps, domain growth slows down and approaches a steady-state value. The plateau values are \(L_1 \approx 200 \Delta x\) 
for spherical particles, \(L_1 \approx 180 \Delta x\) for prolate particles, and \(L_1 \approx 170 \Delta x\) for oblate particles.

In contrast, an alternative measure of domain size, \(L_d\), obtained from the second moments of the full 3D structure factor, does not exhibit the same 
power-law growth. Its increase is significantly slower, with fitted exponents ranging from 0.12 to 0.20. Moreover, \(L_d\) saturates earlier and reaches a 
plateau near \(L_d \approx 70 \Delta x\) for all particle shapes. Although dynamic scaling permits different length scale measures to vary by a constant 
factor, the deviation of \(L_d\) from both viscous and inertial scaling suggests that it may not serve as a reliable characteristic length scale.

Interpreting the structure factor moments as statistical moments of a probability distribution, the second moment of the 3D structure factor corresponds to 
the ratio of the fourth and second moments of the spherically averaged structure factor—due to the \(k^2\) weighting in the volume integral. Thus, \(L_d\) 
may be more appropriately viewed as a descriptor of the structure factor's shape (akin to kurtosis), rather than as a direct measure of domain size.
Nonetheless, the directional second moments \(L_\beta\) provide independent domain size estimates along each Cartesian axis, making them valuable for 
identifying anisotropy in the domain morphology.

\subsection{Effect of magnetic field on bijel formation}

    \begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../figures/results/paper1/D2a-vs-B_ss.png}
    \caption{Dependence of the average domain size on the magnetic field strength $\bar{B}$. The left plot shows the average domain size $L_1$ obtained from the spherically averaged structure factor. The right plot shows the parallel and perpendicular domain size obtained from the respective second moments of the 3D structure factor. Errorbars indicate the standard deviation taken over three independent simulation runs. For anisotropic particles, the directional domain size becomes more anisotropic with increasing field strength.}
    \label{fig:D2a_B}
    \end{figure}

To understand the microstructure changes imparted onto the bijel, we plot the domain size \(L_d\) at the final timestep in 
Figure~\ref{fig:D2a_B}a, normalized by \(R\), the larger semi-axis of the ellipsoidal particles. Each point 
represents the final domain size, averaged over three simulations with identical parameters. This normalization highlights that ellipsoidal particles 
reduce the domain size in bijels, consistent with findings by Günther et al.~\cite{gunther_timescales_2014}. Ellipsoids stabilize a larger interfacial 
area relative to their volume due to their greater cross-sectional area. Additionally, steric effects limit how densely anisotropic particles can pack, 
a topic we discuss further below.

Figure~\ref{fig:D2a_B}b displays the domain size components parallel and perpendicular to the applied magnetic field \(\vec{B}\). For prolate particles, 
the domain size increases by up to three particle radii at the highest field while the perpendicular size remains unchanged. The 
trend reverses for oblate particles, where the parallel domain size slightly decreases and the perpendicular size increases by about four particle radii.
Without a magnetic field, prolate particles yield smaller domains than spherical or oblate ones. However, under field influence, they increase domain size. 
This anisotropy results from particle alignment under the magnetic field. The difference between prolate and oblate behavior 
stems from the dipole orientation. \(\vec{m}\) aligns with the major axis in prolates (within the larger cross-section plane) and the minor axis in oblates 
(perpendicular to it). Thus, field-induced alignment encourages tighter packing perpendicular to \(\vec{B}\) for prolates and parallel for oblates. These 
effects saturate at intermediate field strengths, with little change at higher values.

@@

    \begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{../figures/results/paper1/tortuosity_compare.png}
    \caption{Dependence of the tortuosity on the magnetic field strength $\bar{B}$ for different particle shapes $\alpha$. Different components of the tortuosity tensor show the anisotropy for oblate and prolate particles.}
    \label{fig:tau_B}
    \end{figure*}

The observed anisotropy in domain size suggests that magnetic fields influence bijel morphology at the microscale. To determine whether these structural 
changes impact macroscopic transport properties, we also analyzed the tortuosity of the bijels. While geometric, 
diffusional, and hydraulic definitions of the tortuosity exist, we use the diffusional tortuosity, defined as \(\tau = \epsilon D / D_{\text{eff}}\), 
where \(D_{\text{eff}}\) is the effective diffusivity and \(D\) is the intrinsic molecular diffusivity. \cite{dasilva_tortuosity_2022} Porous structures were 
extracted by binarizing the final order parameter field using a zero threshold, isolating the largest connected (percolating) domain. We computed tortuosity 
along each Cartesian direction using the \texttt{taufactor} package \cite{cooper_taufactor_2016}, which compares steady-state diffusion through the porous 
structure to bulk diffusion in an equivalent volume. Periodic boundary conditions were applied via the \texttt{PeriodicSolver}.

Figure~\ref{fig:tau_B} presents tortuosity as a function of magnetic flux density for the three particle aspect ratios. In the absence of a field, all 
shapes yield a similar tortuosity around \(\tau \approx 1.5\), slightly exceeding the Bruggeman prediction \(\tau = \epsilon^{-0.5} = 1.41\) for equal 
phase volume fractions \(\epsilon = 0.5\) \cite{bruggeman_berechnung_1935, tjaden_origin_2016}. These values align with simulations of gyroid structures 
by Luo et al., who reported tortuosities between 1.48 and 1.73 \cite{luo_macroscopic_2020}.

Under an applied magnetic field, tortuosity becomes directionally dependent for bijels stabilized by anisotropic particles (\(\alpha \neq 1\)). For oblate 
particles (\(\alpha = 0.5\)), tortuosity increases in the field-parallel direction, while remaining near \(\tau \approx 1.5\) perpendicular to the field. 
In contrast, prolate particles (\(\alpha = 2\)) exhibit the opposite trend: tortuosity rises in the perpendicular direction and slightly decreases along 
the field. These trends align with the anisotropic domain sizes and particle alignment under the field. For prolates, alignment of the major axis with the 
magnetic moment results in larger domains and reduced tortuosity along the field direction.

Notably, the most pronounced tortuosity changes occur at intermediate field strength (\(\bar{B} = 0.5\)), suggesting a nonlinear interplay between magnetic 
and interfacial forces. This highlights that field-induced morphological evolution is more complex than a simple monotonic response, prompting a deeper 
investigation into the bijels microstructural changes.
